---
title: "LGH Daily ED visits forecast"
author: "Nayef Ahmad"
date: "2019-10-09"
output: 
    html_document: 
        toc: true
        code_folding: hide
        toc_depth: 4
        toc_float: 
            collapsed: false
        self_contained: no
        keep_md: yes
editor_options: 
  chunk_output_type: console
fontsize: 12
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)

library(knitr)
library(kableExtra)

library(odbc)
library(dbplyr)
library(magrittr)  
library(tidyverse)
library(here)
library(DT)

# shortcuts: 
# > ctrl + alt + i ==> new code chunk
# > ctrl + shift + k ==> knit doc 
# \  ==> line break (include 2 spaces after)

```

# Setup connection to database: 

```{r denodo-setup}
# Setup connections to denodo and import functions

source(here::here("src",
                  "setup-denodo_function.R"))

setup_denodo()


# Functions to pull ED data, and to forecast future ED visits:
source(here::here("src",
                  "ed-visits-denodo_function.R"))
source(here::here("src",
                  "ed-visits-forecast-denodo_function.R"))





```

\  
\  
\  

# Set forecast dates: 

```{r params}
startdate_id <- "20191010"
enddate_id <- "20200103" 

start_date <- ymd(startdate_id)
end_date <- ymd(enddate_id)

data.frame(parameter = c("Beginning of fcast interval", 
                         "End of fcast interval"), 
           date = c(startdate_id, 
                    enddate_id)) %>% 
      
      kable() %>% 
      kable_styling(bootstrap_options = c("striped",
                                          "condensed", 
                                          "responsive"))
                                          


```

\  
\  
\  

# Components of the model: 
```{r fcast-and-actuals}
edvisits_fcast <- edvisits_forecast(startdate_id,
                                    enddate_id,
                                    trend_flexibility = 0.05,
                                    past_years = 5, 
                                    save_plots = TRUE,
                                    holidays_df = holidays)

# No need to pull historical data (it doesn't exist for this interval)
# edvisits_actual <- extract_ed_visits(startdate_id,
#                                      enddate_id)


```

\  
\  
\  

# Historical data and fitted model: 
```{r}
edvisits_fcast[[2]]
```

\  
\  
\  

# Forecast results: 
```{r display-downloadable-table}
edvisits_fcast[[1]] %>%
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip', 
                   buttons = c('excel', "csv"))) %>% 
  formatRound(c("edvisits_fcast", 
                "edvisits_lower_ci", 
                "edvisits_upper_ci"), 
              1)
  
```

\  
\  
\  

# Forecast accuracy
In preliminary validation, we got the following results for 24-day forecast horizon: 

* Cross-validated RMSE is **13.3 visits**. 
* Cross-validated MAE is **10.5 visits**. 
* Cross-validated MAPE is **0.060** or **6.0%**. 

Note that these forecasts are likely to be lower for longer horizons. 




\  
\  
\  

# Forecast plot 



```{r}
# plot comparing actual with forecast: 
edvisits_fcast_df <- edvisits_fcast[[1]]

# plot 
edvisits_fcast_df %>%
      ggplot(aes(x = ds,
                 y = edvisits_fcast)) +
      geom_ribbon(aes(x = ds,
                      ymin = edvisits_lower_ci,
                      ymax = edvisits_upper_ci),
                  fill = "grey80",
                  alpha = 0.5) +
      geom_line(col = "black") +
      geom_point(col = "black") +

      # geom_line(data = edvisits_actual %>% 
      #                 bind_cols(edvisits_fcast_df %>% select(ds)),
      #           aes(x = ds,
      #               y = value),
      #           col = "black" ,
      #           alpha = 0.5) +
      
      labs(title = "LGH ED - Forecast of daily ED visits", 
           subtitle = sprintf("%s to %s \nGrey area represents 95 percent prediction interval", 
                              start_date, 
                              end_date)) + 

      theme_light() +
      theme(panel.grid.minor = element_line(colour = "grey95"),
            panel.grid.major = element_line(colour = "grey95"))

```

\    

# Holiday effects: 


```{r}
edvisits_fcast[[4]] %>% 
  select(ds,
         holidays) %>% 
  filter(ds > ymd(startdate_id)) %>% 
  
  DT::datatable(extensions = 'Buttons', 
                options = list(dom = 'Bfrtip', 
                               buttons = c('excel', "csv"))) 
  
  

```

\  

```{r}
edvisits_fcast[[4]] %>% 
  select(ds,
         holidays) %>% 
  filter(ds > ymd(startdate_id)) %>% 
      
  ggplot(aes(x = ds, 
             y = holidays)) + 
  geom_line() + 
  geom_point() + 
  geom_hline(yintercept = 0, 
             col = "blue") + 
  theme_light() + 
  
  labs(title = "LGH ED - Estimates of holiday effects over forecast interval", 
        subtitle = sprintf("%s to %s",
                           start_date, 
                           end_date), 
       x = "Date", 
       y = "Number of additional ED visits") + 

  theme(panel.grid.minor = element_line(colour = "grey95"), 
        panel.grid.major = element_line(colour = "grey95"))
              


```


\  
\  
\  

# Appendix 

```{r}

# write outputs 

# write_csv(edvisits_fcast_df,
#           here::here("results", 
#                      "output from src",
#                      "2019-10-09_lgh_ED-daily-visits-forecast.csv"))
#                          
# write_csv(edvisits_fcast[[4]] %>% 
#                 select(ds,
#                        holidays) %>% 
#                 filter(ds > ymd(startdate_id)),
#           
#           here::here("results", 
#                      "output from src",
#                      "2019-10-09_lgh_ED-daily-visits-forecast_impacts-of-holidays.csv"))




```

